---
title: "An introduction to `resampling`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{An introduction to `resampling`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 3
)
suppressPackageStartupMessages(library(ComplexHeatmap))
```

## Introduction
**resampling** is an helpful package providing an easy way to take repeated 
samples (i.e. subsets of elements) from a population, here defined as a complete 
set of subjects of interest.

Resampling can be used for different purposes, including the estimation of the 
sampling distribution of an estimator, or the validation of predictive models
in machine learning.

In this article, we show the main functions in `resampling` and a quick example.
For further information, see:

* [Sampling without replacement](sampling/sampling_wor.html)
* [Sampling with replacement](sampling/sampling_wr.html)
* [Resampling without replacement](sampling/resampling_wor.html)
* [Resampling with replacement](sampling/resampling_wr.html)

## Setup
Firstly, we need to load `resampling` and other needed R packages:

```{r setup, message = FALSE}
#resampling
library(resampling)

#Packages for visualisation
require(ComplexHeatmap, quietly = TRUE)
require(grid, quietly = TRUE)
require(RColorBrewer, quietly = TRUE)
```

## Seed
Now we want to set a seed for the random number generation (RNG). In fact, different R sessions have different seeds created from current time and process ID by default, and consequently different simulation results. By fixing a seed we ensure we will be able to reproduce the results of this vignette. We can specify a seed by calling `?set.seed`.


```{r set_seed}
#Set a seed for RNG
set.seed(
  #A seed
  seed = 5381L,                   #a randomly chosen integer value
  #The kind of RNG to use
  kind = "Mersenne-Twister",      #we make explicit the current R default value
  #The kind of Normal generation
  normal.kind = "Inversion"       #we make explicit the current R default value
)
```

<!-- ## Data -->
<!-- We create now some simulated data to use. For this vignette, we will consider 3 -->
<!-- populations: -->

<!-- * A population of 100 elements -->
<!-- * An unbalanced population of 100 elements consisting of 2 subpopulations of 40  -->
<!-- and 60 elements, respectively -->
<!-- * A balanced population of 100 elements consisting of 2 subpopulations of 50 -->
<!-- elements each -->

<!-- ```{r create_data} -->
<!-- #population size -->
<!-- N = 100 -->
<!-- #unbalanced population -->
<!-- unbalanced = c(rep("A", 40), rep("B", 60)) -->
<!-- #balanced population -->
<!-- balanced = c(rep("A", 50), rep("B", 50)) -->
<!-- ``` -->


## Sampling Methods
A list of currently supported sampling methods is available through the 
`?listAvailableSamplingMethods` function call, which returns a table with two 
columns:

* `id`: the id of the sampling method, to be used in the function calls
* `name`: the name of the sampling method


```{r list_sampling_methods}
#list sampling methods
sampling.methods = listAvailableSamplingMethods()

#print in table
knitr::kable(x = sampling.methods)
```


## Sampling Functions
The name of the sampling functions can be retrieved by calling `?listSamplingFunctionNames`.

```{r list_sampling_functions}
#list sampling function names
sampling.function.names = listSamplingFunctionNames()

#print in table
knitr::kable(x = sampling.function.names)
```


Each function is documented. To learn more about a specific method it is possible to use
the `?` operator. For example, let's check the function 
`?balancedSampleWithoutReplacement`.

```{r function_documentation}
#See documentation
?balancedSampleWithoutReplacement
```

From the documentation, we can see that the function accepts three arguments in 
input:

* `strata`: a vector of stratification variables
* `n`: the sample size
* `prob`: an optional vector of probabilities for obtaining the `strata` elements

Let's draw a sample.

```{r example1}
#Balanced sample without replacement
balancedSampleWithoutReplacement(
  strata = c(rep("a", 3),rep("b", 6)),
  n = 6
)
```


## Resampling Functions
The name of the resampling functions can be retrieved by calling `?listResamplingFunctionNames`.

```{r list_resampling_functions}
#list resampling function names
resampling.function.names = listResamplingFunctionNames()

#print in table
knitr::kable(x = resampling.function.names)
```


Each function is documented. To learn more about a specific method it is possible to use
the `?` operator. For example, let's check the function 
`?repeatedStratifiedSampleWithoutReplacement`.

```{r function_documentation2}
#See documentation
?repeatedStratifiedSampleWithoutReplacement
```

From the documentation, we can see that this function takes repeated stratified 
samples without replacement from the population by repeatedly calling the 
`?stratifiedSampleWithoutReplacement` sampling function. 
The `?stratifiedSampleWithoutReplacement` help page reports that the function 
implements the so-called "proportionate allocation", in which the proportion of 
the strata in the population is maintained in the samples.

`?repeatedStratifiedSampleWithoutReplacement` accepts 4 arguments in input:

* `k`: the number of repeated samples to generate
* `strata`: a vector of stratification variables
* `n`: the sample size
* `prob`: an optional vector of probabilities for obtaining the `strata` elements

Let's draw 3 samples.

```{r example2}
#Stratified random samples
repeatedStratifiedSampleWithoutReplacement(
  k = 3,
  strata = c(rep("a", 3),rep("b", 6)),
  n = 6
)
```

We can double-check that the proportion of the strata in the population is 
maintained in the samples:

```{r example3}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Stratified random sample
i = repeatedStratifiedSampleWithoutReplacement(
  k = 2,
  strata = strata,
  n = 3
)

#Check indices
i

#Check ratio
table(strata[i[[1]]])/length(strata[i[[1]]])
table(strata[i[[2]]])/length(strata[i[[2]]])
```


## `resample` Function
Instead of using the different resampling functions, we can use
`?resample` which provides a unique interface to the various resampling methods. 
The parameters in input are:

* `x`: either an integer representing the population size, or a vector of stratification variables
* `n`: either the sample size or the number of elements to holdout
* `k`: the number of repeated samples to generate. It is used as the number of folds in k-fold sampling
* `method`: one of the supported sampling techniques. See `?listAvailableSamplingMethods`
* `prob`: (optional) vector of probability weights for obtaining the elements 
from the population. If provided, its length must match the population size
* `undersample`: logical, whether to remove elements from the population in order to try to obtain balanced data

The function returns an object of class `?resampling`, which represents a series 
of samples repeatedly taken from a population.

## `resampling` S3 Class
The `resampling` class represents a series of samples repeatedly taken from a 
population. A `resampling` object is a `list` consisting of 4 elements:

* `method`: the id of the used sampling method
* `N`: the size of the population from which the samples were taken. Elements in
the population have index from 1 to N
* `removed`: (optional) vector of indices of elements removed from the 
population before taking the samples
* `samples`: list of samples repeatedly taken from the population. Each
element of the list is an integer vector containing the indices of the
elements sampled from the population

Functions to facilitate access to the data stored in a `resampling` object are
available:

* `?getSamplingMethodId`: returns the sampling method id
* `?getPopulationSize`: returns the population size
* `?getRemovedElements`: returns the indices of elements removed from the population
* `?getNumberOfSamples`: returns the number of taken samples
* `?getSamples`: returns the list of samples
* `?getSampleSize`: returns a vector of integer values, the size of each sample
* `?getHoldOutSample`: returns a list where each item is an integer vector 
containing the indices of the elements not sampled from the population
* `?getHoldOutSampleSize`:returns a vector of integer values, the size of each 
hold-out sample

Two other useful functions are `print` and `plot`:

* `?print.resampling`: print a summary of the `resampling` object
* `?plot.resampling`: plot the samples taken from the population as an heatmap

## Take Repeated Samples
Now let's use the `?resample` function to take repeated samples from a population
of 9 elements made of 2 groups (group `a` and group `b`).

```{r setup_strata}
#Define strata
strata = c(rep("a", 3),rep("b", 6))
```

For this example, we want to use **stratified sampling without replacement**. 
If we look at the table returned by `?listAvailableSamplingMethods`, we can see 
that the `id` for stratified sampling without replacement is `stratified_rswor`. 
We can use this value as the `method` argument in `?resample`: 

```{r random_sampling}
#Random sampling without replacement
obj = resample(
  x = strata,
  n = 6,
  k = 2,
  method = "stratified_rswor"
)
```

We can check if the returned object is of class `resampling`:

```{r check_class}
#Is obj of class `resampling`?
is.resampling(obj)
```

Now, we could print a summary:

```{r print_resampling_obj}
#Print
print(obj)
```

We can use `?getSamples` to extract the taken samples from the `resampling` object:

```{r resampling_obj_getsamples}
#Samples
getSamples(obj)
```

The holdout data can be obtained by using `?getHoldOutSamples`:

```{r resampling_obj_getholdoutsamples}
#Holdout Data
getHoldOutSamples(obj)
```

We can also plot our object as an heatmap:

```{r resampling_obj_plot}
#Plot
plot(
  x = obj, 
  strata = strata
)
```


