---
title: "Resampling With Replacement"
# bibliography: references.bib  
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    # toc_float: true
vignette: >
  %\VignetteIndexEntry{Resampling With Replacement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 3
)
```

## Introduction
**Sampling with replacement** implies that each element in the population may appear 
multiple times in one sample. 
There are different techniques of sampling with replacement, including:

* simple random sampling
* random sampling with unequal probabilities
* stratified sampling
* balanced sampling (a special case of stratified sampling)

In this article, we show how to draw **repeated samples with replacement** 
from a population by using the functions implemented in `resampling`.

For further information on the sampling techniques, see 
[Sampling with replacement](sampling_wr.html).

## Setup

### Loading
Firstly, we need to load `resampling` and other needed R packages:

```{r setup, message = FALSE}
#resampling
library(resampling)

#Packages for visualisation
require(ComplexHeatmap, quietly = TRUE)
require(grid, quietly = TRUE)
require(RColorBrewer, quietly = TRUE)
```

### Seed
Then, we set a seed for the random number generation (RNG). In fact, different R 
sessions have different seeds created from current time and process ID by 
default, and consequently different simulation results. 
By fixing a seed we ensure we will be able to reproduce the results of this 
vignette. We can specify a seed by calling `?set.seed`.


```{r set_seed}
#Set a seed for RNG
set.seed(
  #A seed
  seed = 5381L,                   #a randomly chosen integer value
  #The kind of RNG to use
  kind = "Mersenne-Twister",      #we make explicit the current R default value
  #The kind of Normal generation
  normal.kind = "Inversion"       #we make explicit the current R default value
)
```


## Resampling With Replacement
The available methods for taking repeated samples with replacement can be listed 
through the `?listAvailableSamplingMethods` function call, setting the input 
argument to `'rswr'`. 
`?listAvailableSamplingMethods` returns a table with two columns:

* `id`: the id of the sampling method, to be used in the function calls
* `name`: the name of the sampling method

```{r list_sampling_methods}
#list sampling methods
sampling.methods = listAvailableSamplingMethods(x = 'rswr')

#print in table
knitr::kable(x = sampling.methods)
```

The name of the resampling functions can be retrieved by calling `?listResamplingFunctionNames`.

```{r list_resampling_functions}
#list resampling function names
resampling.function.names = listResamplingFunctionNames(x = 'rswr')

#print in table
knitr::kable(x = resampling.function.names)
```


Each function is documented. To learn more about a specific method it is possible to use
the `?` operator. For example, let's check the function 
`?repeatedSimpleRandomSampleWithReplacement`.

```{r function_documentation}
#See documentation
?repeatedSimpleRandomSampleWithReplacement
```

From the documentation, we can see that the function accepts 3 arguments in 
input:

* `k`: the number of repeated samples to generate
* `N`: the population size
* `n`: the sample size


### Simple Random Sampling
In resampling via **simple random sampling** (**SRS**), simple random samples 
with replacement are repeatedly taken from the population.

The function implementing this sampling scheme is 
`?repeatedSimpleRandomSampleWithReplacement`, which accepts 3 arguments:

* `k`: the number of repeated samples to generate
* `N`: the population size
* `n`: the sample size

```{r simple_random_sampling_w_replacement}
#Simple random sampling with replacement
repeatedSimpleRandomSampleWithReplacement(
  k = 2,
  N = 10,
  n = 8
)
```

Instead of using `?repeatedSimpleRandomSampleWithReplacement`, we can take 
repeated samples by using the `?resample` function.

```{r repeated_simple_random_sampling_w_replacement}
#Simple random sampling with replacement
obj = resample(
  x = 10,
  n = 8,
  k = 2,
  method = 'srswr'
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj)
```


### Random Sampling With Unequal Probability
We can use `?repeatedSampleWithReplacement` to draw **repeated samples without replacement with unequal probability**. 
From the documentation, we can see that the function accepts 4 arguments in 
input:

* `k`: the number of repeated samples to generate
* `N`: the population size
* `n`: the sample size
* `prob`: an optional vector of probabilities for obtaining the population elements

For example, let's assume our population of interest has 10 elements, and that 
the first 3 elements have an higher chance of being selected.

```{r random_sampling_w_replacement}
#Random sampling with replacement
repeatedSampleWithReplacement(
  k = 2,
  N = 10,
  n = 5,
  prob = c(rep(3,3), rep(1,7))
)
```

We can take repeated samples by using the `?resample` function and setting 
`method = 'rswr'`.

```{r repeated_random_sampling_w_replacement}
#Random sampling with replacement
obj = resample(
  x = 10,
  n = 5,
  k = 2,
  method = 'rswr',
  prob = c(rep(3,3),rep(1,7))
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj)
```

### Stratified Random Sampling
**Repeated stratified samples with replacement** can be taken by using 
`?repeatedStratifiedSampleWithReplacement` which accept a `strata` argument in input.

```{r stratified_sampling_w_replacement}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Stratified sampling without replacement
repeatedStratifiedSampleWithReplacement(
  k = 2,
  strata = strata,
  n = 6
)
```

We can take repeated samples by using the `?resample` function and setting 
`method = 'stratified_rswr'`.

```{r repeated_stratified_random_sampling_w_replacement}
#Stratified sampling with replacement
obj = resample(
  x = strata,
  n = 6,
  k = 2,
  method = 'stratified_rswr'
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj, strata = strata)
```

### Balanced Random Sampling
**Balanced sampling** is a special case of *stratified sampling* used to ensure that
subgroups of the population are equally represented in the taken sample.

Repeated balanced samples with replacement can be taken by using 
`?repeatedBalancedSampleWithReplacement`:

```{r balanced_sampling_w_replacement}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Balanced sampling with replacement
s = repeatedBalancedSampleWithReplacement(
  k = 2,
  strata = strata,
  n = 6
)

#Check ratio in the samples
table(strata[s[[1]]])/length(strata[s[[1]]])
table(strata[s[[2]]])/length(strata[s[[2]]])
```


We can take repeated samples by using the `?resample` function and setting 
`method = 'balanced_rswr'`.

```{r repeated_balanced_random_sampling_w_replacement}
#Balanced sampling with replacement
obj = resample(
  x = strata,
  n = 6,
  k = 2,
  method = 'balanced_rswr'
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj, strata = strata)
```

### Bootstrap Sampling
We can take **repeated bootstrap samples** by using `?repeatedBootstrapSample`:

```{r bootstrap_sampling}
#Bootstrap sampling
repeatedBootstrapSample(
  k = 2,
  N = 10
)
```


We can also use the `?resample` function by setting `method = 'bootstrap'`.

```{r repeated_bootstrap_sampling}
#Bootstrap sampling
obj = resample(
  x = 10,
  k = 2,
  method = 'bootstrap'
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj)
```




