---
title: "Resampling Without Replacement"
# bibliography: references.bib  
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    # toc_float: true
vignette: >
  %\VignetteIndexEntry{Resampling Without Replacement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 3
)
```

## Introduction
**Sampling without replacement** implies that elements of a population can be 
chosen only once in one sample. 
There are different techniques of sampling without replacement, including:

* simple random sampling
* random sampling with unequal probabilities
* stratified sampling
* balanced sampling (a special case of stratified sampling)
* permutation sampling
* k-fold sampling
* leave-p-out sampling

In this article, we show how to draw **repeated samples without replacement** 
from a population by using the functions implemented in `resampling`.

For further information on the sampling techniques, see 
[Sampling without replacement](sampling_wor.html).

## Setup

### Loading
Firstly, we need to load `resampling` and other needed R packages:

```{r setup, message = FALSE}
#resampling
library(resampling)

#Packages for visualisation
require(ComplexHeatmap, quietly = TRUE)
require(grid, quietly = TRUE)
require(RColorBrewer, quietly = TRUE)
```

### Seed
Then, we set a seed for the random number generation (RNG). In fact, different R 
sessions have different seeds created from current time and process ID by 
default, and consequently different simulation results. 
By fixing a seed we ensure we will be able to reproduce the results of this 
vignette. We can specify a seed by calling `?set.seed`.


```{r set_seed}
#Set a seed for RNG
set.seed(
  #A seed
  seed = 5381L,                   #a randomly chosen integer value
  #The kind of RNG to use
  kind = "Mersenne-Twister",      #we make explicit the current R default value
  #The kind of Normal generation
  normal.kind = "Inversion"       #we make explicit the current R default value
)
```


## Resampling Without Replacement
The available methods for taking repeated samples without replacement can be listed 
through the `?listAvailableSamplingMethods` function call, setting the input 
argument to `'rswor'`. 
`?listAvailableSamplingMethods` returns a table with two columns:

* `id`: the id of the sampling method, to be used in the function calls
* `name`: the name of the sampling method

```{r list_sampling_methods}
#list sampling methods
sampling.methods = listAvailableSamplingMethods(x = 'rswor')

#print in table
knitr::kable(x = sampling.methods)
```

The name of the resampling functions can be retrieved by calling `?listResamplingFunctionNames`.

```{r list_resampling_functions}
#list resampling function names
resampling.function.names = listResamplingFunctionNames(x = 'rswor')

#print in table
knitr::kable(x = resampling.function.names)
```


Each function is documented. To learn more about a specific method it is possible to use
the `?` operator. For example, let's check the function 
`?repeatedSimpleRandomSampleWithoutReplacement`.

```{r function_documentation}
#See documentation
?repeatedSimpleRandomSampleWithoutReplacement
```

From the documentation, we can see that the function accepts 3 arguments in 
input:

* `k`: the number of repeated samples to generate
* `N`: the population size
* `n`: the sample size


### Simple Random Sampling
In resampling via **simple random sampling** (**SRS**), simple random samples 
without replacement are repeatedly taken from the population.

The function implementing this sampling scheme is 
`?repeatedSimpleRandomSampleWithoutReplacement`, which accepts 3 arguments:

* `k`: the number of repeated samples to generate
* `N`: the population size
* `n`: the sample size

```{r simple_random_sampling_wo_replacement}
#Simple random sampling without replacement
repeatedSimpleRandomSampleWithoutReplacement(
  k = 2,
  N = 10,
  n = 8
)
```

Instead of using `?repeatedSimpleRandomSampleWithoutReplacement`, we can take 
repeated samples by using the `?resample` function.

```{r repeated_simple_random_sampling_wo_replacement}
#Simple random sampling without replacement
obj = resample(
  x = 10,
  n = 8,
  k = 2,
  method = 'srswor'
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj)
```


### Random Sampling With Unequal Probability
We can use `?repeatedSampleWithoutReplacement` to draw **repeated samples without 
replacement with unequal probability**. 
From the documentation, we can see that the function accepts 4 arguments in 
input:

* `k`: the number of repeated samples to generate
* `N`: the population size
* `n`: the sample size
* `prob`: an optional vector of probabilities for obtaining the population elements

For example, let's assume our population of interest has 10 elements, and that 
the first 3 elements have an higher chance of being selected.

```{r random_sampling_wo_replacement}
#Random sampling without replacement
repeatedSampleWithoutReplacement(
  k = 2,
  N = 10,
  n = 5,
  prob = c(rep(3,3), rep(1,7))
)
```

We can take repeated samples by using the `?resample` function and setting 
`method = 'rswor'`.

```{r repeated_random_sampling_wo_replacement}
#Random sampling without replacement
obj = resample(
  x = 10,
  n = 5,
  k = 2,
  method = 'rswor',
  prob = c(rep(3,3),rep(1,7))
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj)
```

### Stratified Random Sampling
**Repeated stratified samples without replacement** can be taken by using 
`?repeatedStratifiedSampleWithoutReplacement` which accept a `strata` argument in input.

```{r stratified_sampling_wo_replacement}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Stratified sampling without replacement
repeatedStratifiedSampleWithoutReplacement(
  k = 2,
  strata = strata,
  n = 6
)
```

We can take repeated samples by using the `?resample` function and setting 
`method = 'stratified_rswor'`.

```{r repeated_stratified_random_sampling_wo_replacement}
#Stratified sampling without replacement
obj = resample(
  x = strata,
  n = 6,
  k = 2,
  method = 'stratified_rswor'
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj, strata = strata)
```

### Balanced Random Sampling
**Balanced sampling** is a special case of *stratified sampling* used to ensure that
subgroups of the population are equally represented in the taken sample.

Repeated balanced samples without replacement can be taken by using 
`?repeatedBalancedSampleWithoutReplacement`:

```{r balanced_sampling_wo_replacement}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Balanced sampling without replacement
s = repeatedBalancedSampleWithoutReplacement(
  k = 2,
  strata = strata,
  n = 6
)

#Check ratio in the samples
table(strata[s[[1]]])/length(strata[s[[1]]])
table(strata[s[[2]]])/length(strata[s[[2]]])
```


We can take repeated samples by using the `?resample` function and setting 
`method = 'balanced_rswor'`.

```{r repeated_balanced_random_sampling_wo_replacement}
#Balanced sampling without replacement
obj = resample(
  x = strata,
  n = 6,
  k = 2,
  method = 'balanced_rswor'
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj, strata = strata)
```

### Permutation Sampling
We can take **repeated permutation samples** by using `?repeatedPermutationSample`:

```{r permutation_sampling}
#Permutation sampling
repeatedPermutationSample(
  k = 2,
  N = 10
)
```


We can also use the `?resample` function by setting `method = 'permutation'`.

```{r repeated_permutation_sampling}
#Permutation sampling
obj = resample(
  x = 10,
  k = 2,
  method = 'permutation'
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj)
```

### K-fold Sampling
The idea behind **k-fold resampling** is taken from *k-fold cross-validation*: 
*k* samples are taken from the population, so that the *i*-th sample is 
generated by removing the *i*-th fold from the population and merging 
the remaining *k - 1* folds together.

We can take *k* samples via k-fold sampling by calling `?repeatedRandomKm1Folds`.
The function accepts two arguments in input:

* `k`: the number of folds
* `N`: the population size

`?repeatedRandomKm1Folds` returns a list of length *k* where each element is a 
sample obtained by merging *k-1* folds together.

```{r kfolds_sampling}
#K-1 folds sampling
repeatedRandomKm1Folds(
  N = 10,
  k = 3
)
```

We can take repeated samples by using the `?resample` function and setting
`method = 'kfolds'`.

```{r repeated_kfolds_sampling}
#K-folds sampling
obj = resample(
  x = 10,
  k = 3,
  method = "kfolds"
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(obj)
```

### Stratified K-fold Sampling
We can take repeated stratified samples via **stratified k-fold sampling** by 
calling `?repeatedStratifiedKm1Folds`.

`?repeatedStratifiedKm1Folds` assigns the population to *k* stratified folds and 
returns *k* samples made of elements from *k-1* folds:

```{r stratified_kfolds_sampling}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Assign data to 3 folds
i = repeatedStratifiedKm1Folds(
  strata = strata,
  k = 3
)

#Check folds
i

#Check ratio in the folds
table(strata[i[[1]]])/length(strata[i[[1]]])
table(strata[i[[2]]])/length(strata[i[[2]]])
table(strata[i[[3]]])/length(strata[i[[3]]])
```

We can take repeated samples by using the `?resample` function and setting
`method = 'stratified_kfolds'`.

```{r stratified_kfolds_resampling}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Stratified k-folds sampling
obj = resample(
  x = strata,
  k = 3,
  method = "stratified_kfolds"
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj, strata = strata)
```


### Balanced K-fold Sampling
We can take repeated balanced samples via **balanced k-fold sampling** by 
calling `?repeatedBalancedKm1Folds`, which returns *k* samples made of elements 
from *k-1* folds.

Internally, it uses `?balancedKFolds` to assign the population to *k* balanced folds:

* If the population is balanced, each element in the population is assigned to one 
of the k folds so that the percentage of each stratum is preserved in each fold
* If the population is unbalanced and `undersample = TRUE`, the so-called
*random undersampling* is adopted, i.e. the proportion of the strata in the
population is adjusted by removing elements from the majority groups, so that
each stratum is balanced


```{r balanced_km1folds_sampling}
#Define balanced strata
strata = c(rep(1,6),rep(2,6))

#Check ratio
table(strata)/length(strata)

#Assign data to 3 folds and take a sample
i = repeatedBalancedKm1Folds(
  strata = strata,
  k = 3
)

#Check sample
i

#Check ratio in the sample
table(strata[i[[1]]])/length(strata[i[[1]]])
table(strata[i[[2]]])/length(strata[i[[2]]])
table(strata[i[[3]]])/length(strata[i[[3]]])

#Define unbalanced strata
strata = c(rep(1,6),rep(2,12))

#Check ratio
table(strata)/length(strata)

#Assign data to 3 folds and take a sample
i = repeatedBalancedKm1Folds(
  strata = strata,
  k = 3,
  undersample = T
)
#Check folds
i
#Check ratio in the folds
table(strata[i[[1]]])/length(strata[i[[1]]])
table(strata[i[[2]]])/length(strata[i[[2]]])
table(strata[i[[3]]])/length(strata[i[[3]]])
```


We can take repeated samples by using the `?resample` function.

```{r balanced_kfolds_resampling}
#Define strata
strata = c(rep("a", 6),rep("b", 6))

#Balanced k-folds sampling (balanced population)
obj = resample(
  x = strata,
  k = 3,
  method = "balanced_kfolds"
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj, strata = strata)
```

In case of unbalanced population, we can use the `undersample` argument.

```{r balanced_kfolds_resampling_unbalanced_population}
#Define strata
strata = c(rep("a",6),rep("b",8))

#Balanced k-folds sampling (unbalanced population)
obj = resample(
  x = strata,
  k = 3,
  method = "balanced_kfolds",
  undersample = T
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(x = obj, strata = strata)
```

### Leave-p-out Sampling
**Leave-p-out resampling** is an exhaustive resampling technique, in which samples 
of size `p` are repeatedly taken from the population until all the possible 
combinations of `p` elements are considered. 
These samples are then used as holdout data.

`?repeatedLeavePOut` returns a list of length $\binom{N}{p}$ where 
each element is a sample obtained by removing the holdout data:

```{r leave_p_out_sampling}
#Repeatedly sample leaving out p elements each time
repeatedLeavePOut(N = 5, p = 2)
```

In order to use the `?resample` function, we need to set `method = 'leave_p_out'`.

```{r leave_p_out_resampling}
#Leave-p-out sampling
obj = resample(
  x = 5,
  n = 2,
  method = "leave_p_out"
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(obj)
```

### Leave-one-out Sampling
The *leave-one-out* is a particular case of the *leave-p-out*, where `p = 1`.
Similarly to leave-p-out, **leave-one-out resampling** is an exhaustive 
resampling technique, in which samples of size `1` are repeatedly taken from the 
population until each element in the population is considered as the holdout data.

`?repeatedLeaveOneOut` returns a list of length `N` where 
each element is a sample obtained by removing the holdout data:


```{r leave_one_out_sampling}
#Take one sample leaving out 1 element
repeatedLeaveOneOut(N = 5)
```

In order to use the `?resample` function, we need to set `method = 'leave_one_out'`.

```{r leave_one_out_resampling}
#Leave-one-out sampling
obj = resample(
  x = 5,
  method = "leave_one_out"
)

#Print
print(obj)

#Samples
getSamples(obj)

#Holdout Data
getHoldOutSamples(obj)

#Plot
plot(obj)
```


