---
title: "Sampling Without Replacement"
bibliography: references.bib  
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    # toc_float: true
vignette: >
  %\VignetteIndexEntry{Sampling Without Replacement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 3
)
```

## Introduction
In statistics, *sampling* is the selection of a subset of elements from a 
*population*, here defined as a complete set of subjects of interest. 

Since often it is too expensive or logistically impossible to collect data for 
every case in a population, *sampling* is instead used as a cheap and fast
methodology to estimate its characteristics.

Different sampling schemes exist, but they can be grouped into 2 main categories,
i.e. **sampling with** or **without replacement**: 

* *sampling with replacement* implies each element in the population may appear 
multiple times in one sample
* in *sampling without replacement*, each member of the population can be chosen
only once in one sample

In this article, we show how to draw **samples without replacement** 
from a population by using the functions implemented in `resampling`.

For further information on how to draw *repeated samples without replacement*,
see [Resampling without replacement](resampling_wor.html).

## Setup

### Loading
Firstly, we need to load the `resampling` R package:

```{r setup}
library(resampling)
```

### Seed
Then, we set a seed for the random number generation (RNG). In fact, different R 
sessions have different seeds created from current time and process ID by 
default, and consequently different simulation results. 
By fixing a seed we ensure we will be able to reproduce the results of this 
vignette. We can specify a seed by calling `?set.seed`.


```{r set_seed}
#Set a seed for RNG
set.seed(
  #A seed
  seed = 5381L,                   #a randomly chosen integer value
  #The kind of RNG to use
  kind = "Mersenne-Twister",      #we make explicit the current R default value
  #The kind of Normal generation
  normal.kind = "Inversion"       #we make explicit the current R default value
)
```


## Sampling Without Replacement
As previously wrote, **sampling without replacement** implies that elements of a 
population can be chosen only once in one sample. 
There are different techniques of sampling without replacement, including:

* simple random sampling
* random sampling with unequal probabilities
* stratified sampling
* balanced sampling (a special case of stratified sampling)
* permutation sampling
* k-fold sampling
* leave-p-out sampling

The available methods can be listed through the `?listAvailableSamplingMethods` 
function call, setting the input argument to `'rswor'`. 
`?listAvailableSamplingMethods` returns a table with two columns:

* `id`: the id of the sampling method, to be used in the function calls
* `name`: the name of the sampling method

```{r list_sampling_methods}
#list sampling methods
sampling.methods = listAvailableSamplingMethods(x = 'rswor')

#print in table
knitr::kable(x = sampling.methods)
```

The name of the sampling functions can be retrieved by calling `?listSamplingFunctionNames`.

```{r list_sampling_functions}
#list sampling function names
sampling.function.names = listSamplingFunctionNames(x = 'rswor')

#print in table
knitr::kable(x = sampling.function.names)
```


Each function is documented. To learn more about a specific method it is possible to use
the `?` operator. For example, let's check the function 
`?simpleRandomSampleWithoutReplacement`.

```{r function_documentation}
#See documentation
?simpleRandomSampleWithoutReplacement
```

From the documentation, we can see that the function accepts 2 arguments in 
input:

* `N`: the population size
* `n`: the sample size


### Simple Random Sampling
**Simple random sampling** (**SRS**) is the easiest form of sampling without replacement.
In SRS without replacement, each element of the population has the same probability 
of being selected for the sample.

```{r simple_random_sampling_wo_replacement}
#Simple random sampling without replacement
simpleRandomSampleWithoutReplacement(
  N = 10,
  n = 8
)
```




### Random Sampling With Unequal Probability
The concept of **random sampling without replacement with unequal probability** was 
developed by Narain [@Narain1951], Horvitz and Thompson [@Horvitz1952].
Under this sampling design, elements of the population have different 
probabilities of being selected.
We can use `?sampleWithoutReplacement` to draw our sample. For example, let's 
assume our population of interest has 10 elements, and that the first 3 elements
have an higher chance of being selected.

```{r random_sampling_wo_replacement}
#Random sampling without replacement
sampleWithoutReplacement(
  N = 10,
  n = 5,
  prob = c(rep(3,3), rep(1,7))
)
```



### Stratified Random Sampling
When a population can be partitioned into groups (i.e. *strata* or subpopulations)
having certain properties in common, a **stratified sampling** approach can be used. 
This sampling design is adopted to ensure that subgroups of the population are 
represented in the taken sample.

A stratified sample without replacement can be taken by using 
`?stratifiedSampleWithoutReplacement` which accept a `strata` argument in input.

```{r stratified_sampling_wo_replacement}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Stratified sampling without replacement
stratifiedSampleWithoutReplacement(
  strata = strata,
  n = 9
)
```

`?stratifiedSampleWithoutReplacement` implements the so-called "proportionate 
allocation", in which the proportion of the strata in the population is 
maintained in the samples.

```{r stratified_sampling_wo_replacement_proportionate_allocation}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Stratified sampling without replacement
s = stratifiedSampleWithoutReplacement(
  strata = strata,
  n = 6
)

#Check ratio in the sample
table(strata[s])/length(strata[s])
```



### Balanced Random Sampling
**Balanced sampling** is a special case of *stratified sampling* used to ensure that
subgroups of the population are equally represented in the taken sample.

```{r balanced_sampling_wo_replacement}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Balanced sampling without replacement
s = balancedSampleWithoutReplacement(
  strata = strata,
  n = 6
)

#Check ratio in the sample
table(strata[s])/length(strata[s])
```



### Permutation Sampling
A **permutation sample** is a sample of the same size of the population, where the
elements are simply rearranged in a random order.

A permutation sample can be taken by using `?permutationSample`:

```{r permutation_sampling}
#Permutation sampling
permutationSample(
  N = 10
)
```



### K-fold Sampling
In **k-fold sampling**, each element of the population is randomly assigned to 1 
of *k* folds. `resampling` provides a function to easily select a random sample
made of elements from *k-1* folds:

```{r kfolds_sampling}
#K-1 folds sampling
randomKm1Folds(
  N = 10,
  k = 3
)
```


### Stratified K-fold Sampling
In **stratified k-fold sampling**, each element in the population is assigned to 
one of the *k* folds so that the percentage of each stratum in the population is 
preserved in each fold.

`?stratifiedKm1Folds` assigns the population to *k* stratified folds and returns
a sample made of elements from *k-1* folds:

```{r stratified_kfolds_sampling}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Assign data to 3 folds and take a sample
i = stratifiedKm1Folds(
  strata = strata,
  k = 3
)

#Check folds
i

#Check ratio in the folds
table(strata[i])/length(strata[i])
```



### Balanced K-fold Sampling
**Balanced k-fold sampling** is a special case of *stratified sampling* in which
the population is assigned to *k* balanced folds.

`?balancedKm1Folds` assigns the population to *k* balanced folds and returns
a sample made of elements from *k-1* folds. 
Internally, it uses `?balancedKFolds` to assign the population to *k* balanced folds:

* If the population is balanced, each element in the population is assigned to one 
of the k folds so that the percentage of each stratum is preserved in each fold
* If the population is unbalanced and `undersample = TRUE`, the so-called
*random undersampling* is adopted, i.e. the proportion of the strata in the
population is adjusted by removing elements from the majority groups, so that
each stratum is balanced


```{r balanced_km1folds_sampling}
#Define balanced strata
strata = c(rep(1,6),rep(2,6))

#Check ratio
table(strata)/length(strata)

#Assign data to 3 folds and take a sample
i = balancedKm1Folds(
  strata = strata,
  k = 3
)

#Check sample
i

#Check ratio in the sample
table(strata[i])/length(strata[i])

#Define unbalanced strata
strata = c(rep(1,6),rep(2,12))

#Check ratio
table(strata)/length(strata)

#Assign data to 3 folds and take a sample
i = balancedKm1Folds(
  strata = strata,
  k = 3,
  undersample = T
)

#Check folds
i

#Check ratio in the folds
table(strata[i])/length(strata[i])
```




### Leave-p-out Sampling
In **leave-p-out** sampling, a random sample of size `p` is taken from the 
population and used as holdout data.

`?leavePOutSample` returns a random sample obtained by removing the holdout
sample from the population:

```{r leave_p_out_sampling}
#Take one sample leaving out p elements
leavePOutSample(N = 5, p = 2)
```



### Leave-one-out Sampling
The **leave-one-out** is a particular case of the *leave-p-out*, where `p = 1`.
Particularly, a random sample of size `1` is taken from the population and used 
as holdout data.

For example, let's assume our population of interest has 5 elements. In this 
case, a sample obtained with this sampling technique will result in 4 elements:

```{r leave_one_out_sampling}
#Take one sample leaving out 1 element
leaveOneOutSample(N = 5)
```




## References

