---
title: "Sampling With Replacement"
bibliography: references.bib  
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    # toc_float: true
vignette: >
  %\VignetteIndexEntry{Sampling With Replacement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 3
)
```

## Introduction
In statistics, *sampling* is the selection of a subset of elements from a 
*population*, here defined as a complete set of subjects of interest. 

Since often it is too expensive or logistically impossible to collect data for 
every case in a population, *sampling* is instead used as a cheap and fast
methodology to estimate its characteristics.

Different sampling schemes exist, but they can be grouped into 2 main categories,
i.e. **sampling with** or **without replacement**: 

* *sampling with replacement* implies each element in the population may appear 
multiple times in one sample
* in *sampling without replacement*, each member of the population can be chosen
only once in one sample

In this article, we show how to draw **samples with replacement** 
from a population by using the functions implemented in `resampling`.

For further information on how to draw *repeated samples with replacement*,
see [Resampling with replacement](resampling_wr.html).

## Setup

### Loading
Firstly, we need to load the `resampling` R package:

```{r setup}
library(resampling)
```

### Seed
Then, we set a seed for the random number generation (RNG). In fact, different R 
sessions have different seeds created from current time and process ID by 
default, and consequently different simulation results. 
By fixing a seed we ensure we will be able to reproduce the results of this 
vignette. We can specify a seed by calling `?set.seed`.


```{r set_seed}
#Set a seed for RNG
set.seed(
  #A seed
  seed = 5381L,                   #a randomly chosen integer value
  #The kind of RNG to use
  kind = "Mersenne-Twister",      #we make explicit the current R default value
  #The kind of Normal generation
  normal.kind = "Inversion"       #we make explicit the current R default value
)
```


## Sampling With Replacement
As previously wrote, **sampling with replacement** implies that elements of a 
population can be chosen multiple times in one sample. 
There are different techniques of sampling with replacement, including:

* simple random sampling
* random sampling with unequal probabilities
* stratified sampling
* balanced sampling (a special case of stratified sampling)

The available methods can be listed through the `?listAvailableSamplingMethods` 
function call, setting the input argument to `'rswr'`. 
`?listAvailableSamplingMethods` returns a table with two columns:

* `id`: the id of the sampling method, to be used in the function calls
* `name`: the name of the sampling method

```{r list_sampling_methods}
#list sampling methods
sampling.methods = listAvailableSamplingMethods(x = 'rswr')

#print in table
knitr::kable(x = sampling.methods)
```

The name of the sampling functions can be retrieved by calling `?listSamplingFunctionNames`.

```{r list_sampling_functions}
#list sampling function names
sampling.function.names = listSamplingFunctionNames(x = 'rswr')

#print in table
knitr::kable(x = sampling.function.names)
```


Each function is documented. To learn more about a specific method it is possible to use
the `?` operator. For example, let's check the function 
`?simpleRandomSampleWithReplacement`.

```{r function_documentation}
#See documentation
?simpleRandomSampleWithReplacement
```

From the documentation, we can see that the function accepts 2 arguments in 
input:

* `N`: the population size
* `n`: the sample size

### Simple Random Sampling
**Simple random sampling** (**SRS**) is the easiest form of sampling with replacement.
In SRS with replacement, each element of the population has the same probability 
of being selected for the sample.

```{r simple_random_sampling_w_replacement}
#Simple random sampling with replacement
simpleRandomSampleWithReplacement(
  N = 10,
  n = 8
)
```


### Random Sampling With Unequal Probability
The concept of **random sampling with unequal probability** was perhaps introduced by 
Hansen and Hurwitz in the context of sampling with replacement [@Hansen1943]. 
Under this sampling design, elements of the population have different 
probabilities of being selected.
We can use `?sampleWithReplacement` to draw our sample. For example, let's 
assume our population of interest has 10 elements, and that the first 3 elements
have an higher chance of being selected.

```{r random_sampling_w_replacement}
#Random sampling with replacement
sampleWithReplacement(
  N = 10,
  n = 8,
  prob = c(rep(3,3), rep(1,7))
)
```



### Stratified Random Sampling
When a population can be partitioned into groups (i.e. *strata* or subpopulations)
having certain properties in common, a **stratified sampling** approach can be used. 
This sampling design is adopted to ensure that subgroups of the population are 
represented in the taken sample.

A **stratified sample with replacement** can be taken by using 
`?stratifiedSampleWithReplacement` which accept a `strata` argument in input.

```{r stratified_sampling_w_replacement}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Stratified sampling with replacement
stratifiedSampleWithReplacement(
  strata = strata,
  n = 9
)
```

`?stratifiedSampleWithReplacement` implements the so-called "proportionate 
allocation", in which the proportion of the strata in the population is 
maintained in the samples.

```{r stratified_sampling_w_replacement_proportionate_allocation}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Stratified sampling with replacement
s = stratifiedSampleWithReplacement(
  strata = strata,
  n = 9
)

#Check ratio in the sample
table(strata[s])/length(strata[s])
```



### Balanced Random Sampling
**Balanced sampling** is a special case of *stratified sampling* used to ensure that
subgroups of the population are equally represented in the taken sample.

```{r balanced_sampling_w_replacement}
#Define strata
strata = c(rep("a", 3),rep("b", 6))

#Check ratio
table(strata)/length(strata)

#Balanced sampling with replacement
s = balancedSampleWithReplacement(
  strata = strata,
  n = 8
)

#Check ratio in the sample
table(strata[s])/length(strata[s])
```




### Bootstrap Sampling
Inspired by earlier work on the jackknife, Bradley Efron published the bootstrap
method in 1979 [@Efron1979]. An **ordinary bootstrap sample** is a special case 
of random sample with replacement where the sample size is equivalent to the 
population size.

A bootstrap sample can be taken by using `?bootstrapSample`:

```{r bootstrap_sampling}
#Bootstrap sampling
bootstrapSample(
  N = 10
)
```



## References

